%=========================================================================

\chapter{Rdrand-Gen} \label{chap:generator}

Because the library is only for C~language, using it for example with shell scripts would be difficult. For this reason was created also a~simple executable application, which is installed with the~library, and which can be used for generating random values without the~need of using C~language.

The generator has four optional command-line parameters to~modify its behavior. Firstly, {\tt --amount} can be used to generate specific amount of~bytes of~randomness. Suffixes K, M, G and T are accepted for easier use and when this option is not used, then the~application is generating indefinitely until it is stopped, for example by KILL signal.

The second parameter is {\tt --method}, which allows user to change the default method \\\function{rdrand_get_bytes_retry} for the~two reseeding functions. The~names of the~methods are made shorter for the~interaction with the~user. Third parameter {\tt --output} is used for specifying the output file -- without it, the random values are printed on {\tt stdout}. 

The {\tt --threads} can specify, how many threads the~generator will run in parallel for better performance. By default it is set to 2, because according of Intel~\cite{IntelArk}, in Ivy Bridge generation of CPUs, in which the~instruction was added, there are still CPUs with just two processing units (i.e. the Celeron serie).

\section{Underflow Recovery}
Although is stated in Intel's Software Developer Manual~\cite[chapter~7.3.17]{IntelSWManualVol1}, that an exceeding of the speed of the internal generator is unlikely, and according of {\em unverified} information, for example on StackOverflow~\cite{StackoverflowRDRANDCharacteristics} it should not be possible in current generations (specifically on Ivy Bridge) of Intel's CPU to achieve it, we decided that the application should be working with good performance even in case of slower internal generator. The importance of this decision become even more obvious after finding that on \machine{dell-pr1700-02.lab.bos.redhat.com} the CPU wasn't able to handle more than four parallel threads reading from RdRand\footnote{Unfortunately, I cannot provide a statistic probability of such situation -- only one machine from all I have tried had this problem and I wasn't able to find that anyone other would experienced this too, which is, with regards of the currently low usage of the instruction, not surprising.}.

The principle is simple: By default, there is tolerance for few failures, implemented in the library itself, when a new call of the RdRand instruction is make immediately. But if amount of the failures in a row exceeds a limit with which the generator application was compiled, then there is a need for another approach; it is necessary to be able to work even on CPU with slow generator, yet still be able to detect complete failure of the HW.

In the case of exceeding of the HW RNG speed, the generator application tries to lower its own speed to get aligned with HW. This is at first done by decrementing threads count by one and new try. If this solution is not working or is not possible (that means, when the threads count was lowered to a single thread, or was so from the beginning), delays are being inserted between calls. The delays are then lengthened with each unsuccessful call. If even in this case the HW RNG is not able to provide enough random values, the application ends with an error message\footnote{Such situation would be clearly a sign of a hardware error and thus it is questionable if the generated values would be still really random}.

%=========================================================================
